Soling at the 2000 Summer Olympics

### 2020-07-27 inst/f1points.Rmd
more reconfiguration

@@

The points scored by each driver give a ranking shown in the table;
but Plackett-Luce likelihoods also give a ranking that does not depend
on any arbitrary points system.  We can plot one ranking against the
other, see Figure \ref{orderingbypointsandlikelihood}\footnote{The
historically correct points awarded to the drivers differs from that
calculated here.  That is for two reasons: firstly, ``fastest lap''
points are not included here, and also the truncation of the order
table to the first 11 drivers can increase the rank of a driver if
non-first-11 drivers are placed}.  We thus see a comparison between
the two ordering systems.  Taking 2017 as an example, drivers Hamilton
and Vettel are are respectively first and second according to both
ranking ranking procedures; but R\"{a}ikk\"{o}nen YES I KNOW IT IS
WRONG: RMARKDOWN IS DEFECTIVE and Bottas are third and fourth, and
fourth and third, according to points and likelihood respectively.
The degree of agreement between the two systems might be quantified as
"2", as the positions of the top two drivers agree.  Thus the degree
of agreement between likelihood ranking and points ranking for the
years 2016-9 would be 8,2,6, and 2 respectively.

@@

```{r, label=calcpoints_zipf, echo=FALSE, cache=TRUE}
p <- formula1_points_systems()$zipf
dp2016 <- ordertable2points(as.ordertable(F1_table_2016[seq_len(top),]),p)
dp2016 <- sort(dp2016,decreasing=TRUE)
dp2016[] <- seq_along(dp2016)
m2016 <- maxp(H2016)
m2016 <- sort(m2016,decreasing=TRUE)
m2016[] <- seq_along(m2016)

@@

## Points systems

Here we follow the above protocol to assess different points systems.
There are a number of plausible points systems that might be used:

@@

Applying the current points system, for example, to the 2017 table we
would rank the drivers as follows:

```{r,label=realpointsshower,echo=FALSE}
real <- ordertable2points(ranktable_to_ordertable(wikitable_to_ranktable(F1_table_2017)),p)
real <- round(real[seq_len(top)],2)
real <- sort(real,decreasing=TRUE)
real[] <- seq_along(real)
real
```

```{r,label=zipfpointsshower,echo=FALSE}
zipf <- ordertable2points(ranktable_to_ordertable(wikitable_to_ranktable(F1_table_2017)),1)
zipf <- round(zipf[seq_len(top)],2)
zipf <- sort(zipf,decreasing=TRUE)
zipf[] <- seq_along(zipf)
zipf
```

>>>>

Applying the current points system, for example, to the 2017 results table we
would rank the drivers as follows:

```{r,label=realpointsshower,echo=FALSE}
real <- ordertable2points(ranktable_to_ordertable(wikitable_to_ranktable(F1_table_2017)),formula1_points_systems(top)$real)
real <- round(real[seq_len(top)],2)
real <- sort(real,decreasing=TRUE)
real[] <- seq_along(real)
real
```

```{r,label=zipfpointsshower,echo=FALSE}
zipf <- ordertable2points(ranktable_to_ordertable(wikitable_to_ranktable(F1_table_2017)),formula1_points_systems(top)$zipf)
zipf <- round(zipf[seq_len(top)],2)
zipf <- sort(zipf,decreasing=TRUE)
zipf[] <- seq_along(zipf)
zipf
```

@@

We can plot one ranking against the other\footnote{The historically
correct points awarded to the drivers differs from that calculated
here.  That is for two reasons: firstly, ``fastest lap'' points are
not included here, and also the truncation of the order table to the
first 11 drivers can increase the rank of a driver if non-first-11
drivers are placed} , shown in Figure
\ref{orderingbypointsandlikelihood}.  We thus see a comparison between
the two ordering systems.  Taking 2017 as an example, drivers Hamilton
and Vettel are are respectively first and second according to both
ranking ranking procedures; but R\"{a}ikk\"{o}nen YES I KNOW IT IS
WRONG: RMARKDOWN IS DEFECTIVE and Bottas are third and fourth, and
fourth and third, according to points and likelihood respectively.
The degree of agreement between the two systems might be quantified as
"2", as the positions of the top two drivers agree.  Thus the degree
of agreement between likelihood ranking and points ranking for the
years 2016-9 would be 8,2,6, and 2 respectively.

@@

```{r, label=calcpoints_zipf, echo=FALSE, cache=TRUE}
pz <- formula1_points_systems()$zipf
dp2016z <- ordertable2points(as.ordertable(F1_table_2016[seq_len(top),]),pz)
dp2016z <- sort(dp2016z,decreasing=TRUE)
dp2016z[] <- seq_along(dp2016z)

dp2017z <- ordertable2points(as.ordertable(F1_table_2017[seq_len(top),]),pz)
dp2017z <- sort(dp2017z,decreasing=TRUE)
dp2017z[] <- seq_along(dp2017z)

dp2018z <- ordertable2points(as.ordertable(F1_table_2018[seq_len(top),]),pz)
dp2018z <- sort(dp2018z,decreasing=TRUE)
dp2018z[] <- seq_along(dp2018z)

dp2019z <- ordertable2points(as.ordertable(F1_table_2019[seq_len(top),]),pz)
dp2019z <- sort(dp2019z,decreasing=TRUE)
dp2019z[] <- seq_along(dp2019z)
```

@@

par(pty='s')        # square plot
ox <- dp2016z
oy <- ordertrans(m2016,dp2016z)
plot(ox,oy,asp=1,pty='s',xlim=c(0,11.2),ylim=c(0,11.2),pch=16,
    xlab="points order",ylab="likelihood order",main='2016 season')
par(xpd=TRUE)       # allow drivers' names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(ox)[i],pos=4,col='gray',cex=0.6) }
par(xpd=FALSE)      # stop diagonal line from protruding beyond plotting region
abline(0,1)

par(pty='s')        # square plot
ox <- dp2017z
oy <- ordertrans(m2017,dp2017z)
plot(ox,oy,asp=1,pty='s',xlim=c(0,11.2),ylim=c(0,11.2),pch=16,
    xlab="points order",ylab="likelihood order",main='2017 season')
par(xpd=TRUE)       # allow drivers' names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(ox)[i],pos=4,col='gray', cex=0.6) }
par(xpd=FALSE)      # stop diagonal line from protruding beyond plotting region
abline(0,1)

par(pty='s')        # square plot
ox <- dp2018z
oy <- ordertrans(m2018,dp2018z)
plot(ox,oy,asp=1,pty='s',xlim=c(0,11.2),ylim=c(0,11.2),pch=16,
    xlab="points order",ylab="likelihood order",main='2018 season')
par(xpd=TRUE)       # allow drivers' names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(ox)[i],pos=4,col='gray', cex=0.6) }
par(xpd=FALSE)      # stop diagonal line from protruding beyond plotting region
abline(0,1)

par(pty='s')        # square plot
ox <- dp2019z
oy <- ordertrans(m2019,dp2019z)
plot(ox,oy,asp=1,pty='s',xlim=c(0,11.2),ylim=c(0,11.2),pch=16,
    xlab="points order",ylab="likelihood order",main='2019 season')
par(xpd=TRUE)       # allow drivers' names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(ox)[i],pos=4,col='gray',cex=0.6) }
par(xpd=FALSE)      # stop diagonal line from protruding beyond plotting region
abline(0,1)
```

## Possible points systems

There are a number of plausible points systems that might be used:

@@
